Real-time resolving assembly graph by ONT long reads data

### 2019-11-19 paper.tex
bmc

Nonetheless, the challenges to adapt this approach into a real-time mechanism are obvious, mainly from building a progressive, self-updating path-finding and graph reducing method which are required for a streaming assembler on graph. 
A modified DFS (depth-first search) algorithm and a voting system with comprehensive scoringn system has been implemented to overcome these issues.
This results in \npgraph{}, a user-friendly tool with GUI that can traverse the assembly graph and bridge its components in real-time as long as the nanopore sequencing process is still running.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{\npgraph{} -- a tool to resolve assembly graph in real-time}
\npgraph{}'s input consists of Illumina assembly graph resulted from running assembler, e.g. \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$ \cite{Zerbino2008}, $\mathtt{AbySS}$ \cite{Simpson2009} on Illumina short reads, together with long reads from third generation sequencing technology (Oxford Nanopore Technology, Pacbio).
The long reads will be aligned with the contigs in the assembly graph to indicate longer paths that should be traversed. These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
\npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

Compared to \npscarf, \npgraph{} employs less greedy approach based on graph traversal. This might reduce the bridging rate when the linking conditions become more strict, but the concurrent results reported are more reliable.
The performance of \npgraph{} is comparable to \unicycler{} while consuming much less computational resources so that it can work on streaming mode. 
In addition, the integrated GUI allows users to interact in more efficient way.

As a hybrid assembler, similar to \unicycler{}, \npgraph{} relies on the initial assembly graph to generate the final assembly. The algorithm operates on the assumption of a high quality assembly from sufficient Illumina data that produces decent assembly graph, with a minimal amount of long reads data possible. A fragmented initial graph directly affects the quality of the final assembly, independant from the abundance of nanopore data. This issue has been reflected via our attempt to apply \npgraph{} on the zymo metagenomics data.

@@

It is worth noting that due to dead-ends, there not always exist a path in the assembly graph connecting two anchors as suggested by the alignments. In this case, if enough long reads coverage (20X) are met, a consensus module is invoked and the resulting sequence is contained in a \emph{pseudo} edge.
\subsection*{Graph simplification in real-time}
\npgraph{} resolves the graph by reducing its complexity perpetually using the long reads that can be provided in real-time.
Whenever a bridge is finished (with a unique linking path), the assembly graph is \emph{transformed} or \emph{reduced} by replacing its unique path with a composite edge and removing any unique edges (edges coming from unique nodes) along the path. The assembly graph would have at least one edge less than the original after the reduction. The nodes located on the reduced path, other than 2 ends, also have their multiplicities subtracted by one and the bridge is marked as finally resolved without any further modifications.

>>>>

Nonetheless, the challenges to adapt this approach into a real-time mechanism are obvious, mainly from building a progressive, self-updating path-finding and graph reducing method which are required for a streaming assembler on graph. 
A modified DFS (depth-first search) algorithm and a progressive voting algorithm has been implemented to overcome these issues.
This results in \npgraph{}, a user-friendly tool with GUI that can traverse the assembly graph and bridge its components in real-time as long as the nanopore sequencing process is still running.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{\npgraph{} -- a tool to resolve assembly graph in real-time}
\npgraph{}'s input consists of Illumina assembly graph resulted from running assembler, e.g. \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$ \cite{Zerbino2008}, $\mathtt{AbySS}$ \cite{Simpson2009} on Illumina short reads, together with long reads from third generation sequencing technology (Oxford Nanopore Technology, Pacbio).
The long reads will be aligned with the contigs in the assembly graph to indicate longer paths that should be traversed. These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
\npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

Compared to \npscarf, \npgraph{} algorithm employs a less greedy approach based on graph traversal. This might reduce the bridging rate when the linking conditions become more strict, but the concurrent results reported are more reliable.
The performance of \npgraph{} is comparable to \unicycler{} while consuming much less computational resources so that it can work on streaming mode. 
Also, the integrated GUI allows users to visualize its animated output in a  more efficient way.

In the very first attempt to run real-time assembly on metagenomics data, \npgraph{} proved that its method can adapt to scaled-up data and complexity. The assembler was able to handle the De Bruijin assembly graph from Illumina data generated by $\mathtt{metaSPAdes}$ and simplify it using a minimal input stream of long reads. 
The result has shown more complete isolate genomes reconstructed out of the graph, but the overall communities` genome coverage were dropped with several strains omitted comparing to the reference. The issue is reasoned by the insufficient short-read input to cover underpresented species in the mixture, especially in the case of Log community.
As a hybrid assembler, similar to \unicycler{}, \npgraph{} relies on the initial assembly graph to generate the final assembly. The algorithm operates on the assumption of a high quality assembly from sufficient Illumina data that produces decent assembly graph, with a minimal amount of long reads data possible. A fragmented initial graph directly affects the quality of the final assembly, free from the abundance of nanopore data.

@@

It is worth noting that due to dead-ends, there not always exist a path in the assembly graph connecting two anchors as suggested by the alignments. In this case, if enough long reads coverage (20X) are met, a consensus module is invoked and the resulting sequence is contained in a \emph{pseudo} edge.
\subsection*{Graph simplification in real-time}
\npgraph{} resolves the graph by reducing its complexity perpetually using the long reads that can be streamed in real-time.
Whenever a bridge is finished (with a unique linking path), the assembly graph is \emph{transformed} or \emph{reduced} by replacing its unique path with a composite edge and removing any unique edges (edges coming from unique nodes) along the path. The assembly graph would have at least one edge less than the original after the reduction. The nodes located on the reduced path, other than 2 ends, also have their multiplicities subtracted by one and the bridge is marked as finally resolved without any further modifications.
